Learning Which Bad Smells are Real Stinkers?

### 2016-02-28 Manuscript.tex
Updates from ShareLaTeX

\usepackage[shortlabels]{enumitem} 
\usepackage{amsmath}
\usepackage{url}
\usepackage{balance}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
%\setlist{nolistsep,leftmargin=5mm}
%\usepackage[pdftex]{graphicx}
\usepackage{program}
\newcommand{\Sample}{{\bf SAMPLE}}
\newcommand{\PEEKING}{{\bf PEEKING2}}
%\usepackage[table]{xcolor}
\definecolor{darkgreen}{rgb}{0,0.3,0}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\usepackage{colortbl}
\usepackage{picture}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{listings}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

\title{How to Learn Useful Changes to Software Projects\\(to Assist Refactoring)}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Tennessee, USA}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract}

@@

When modern software systems are developed, their design changes continuously. Due to market constraints, the developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}). These anti-patterns are generally viewed as poor solutions to recurring design problems, possibly deterring the maintainability of the system. In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{}\cite{}\cite{}. Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{}\cite{}. This makes early detection of these code smells a necessity to maintain sufficient quality in software systems. This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed by (i) identifying potential anti-patterns; (ii) generating reliable refactoring options, and (iii) critiquing bad refactoring options.

Tools for detecting code smells have been the focus of several researchers lately, \cite{}. One popular analysis technique is the use of static code analysis and code quality metrics\cite{}. While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers in planning actions to improve the quality of source code.  Similar to the concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.

At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. The panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. We call such tools \textit{planners}. In the context of detecting code smells, the planners are designed to address the following issues.

@@

In response to this, we propose a novel {\em planning} method called XTREE which reflects on static code quality metrics to learn refactoring options to a software system such that its performance ``improves'' quantitatively. As a measure of source code quality we take into account the number of classes with defects in them. To this end, we use defect count data set by Jureczko et al. for JAVA systems~\cite{jureczko10}.

@@

\bi
\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reduce the number of defects. Specifically, we use XTREE to suggest valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}.
\item \textbf{RQ2:} \textit{Are the plans suggested by a planner succinct?} The second question aims to measure the number of refactorings required 
\item \textbf{RQ3:} \textit{Insightful: Do the plans tell us thing we do not expect?}
\ei

@@

\section{Preliminaries}\label{sect:prelim}

\subsection{Code Smells?}

Anti-patterns are poor solutions and recurring design problems that are unwillingly introduced by developers while designing and implementing large object-oriented systems. The impact of these code smells have been subjected to several empirical studies. As a result of the, the identification of these anti patterns has been receiving more attention from both researchers and practitioners.

During lifecycle, a software system undergoes continuous changes in order to keep up with the 
business level demands \cite{36}. Unfortunately, due to market/customers constraints, the source code quality is often neglected with the risk of introducing \textit{code bad smell} (or code-smells) \cite{24}.

Fowler \cite{24} and Brown et al. \cite{13} defined a catalogue of more than 30 anti-patterns. For each anti-pattern, they reported the definition of the anti-pattern as well as specific refactoring operations aimed at removing it in order to improve the quality of source code.

To illustrate this, consider for example developers adding several responsibilities to a class rather that including them in separate classes. This results in a class that grows rapidly, and when there are further added responsibilities, the classes grow and breed. Ultimately, the class becomes too complex and its quality deteriorates. Such classes are know as Large Class \cite{24} and could be the cause of a design problem in source code.

For a while now, there have been anecdotal evidence that design problems (such as anti-patterns) negatively impacts software comprehension and maintenance. However, in the past few years, anti-patterns have been the subject of several empirical studies aiming at empirically analyzing their impact on software maintainability. From these studies, we now have empirical evidence that code containing code smells or participating in anti-patterns is significantly more change-prone than ``clean'' code \cite{31}. Also, code participating in anti-patterns has a higher fault-proneness than the rest of the system code \cite{31}\cite{37}.

Other studies aimed at understanding the impact of anti-patterns on program comprehension \cite{1} showed that the presence of an anti-pattern in the source code does not decrease the developers' performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{1}\cite{61}. While the results of this study indicates that single anti-patterns are not harmful, they also reveals that anti-patterns are quite diffuse in software systems and very often code components are affected by more than one anti-pattern. In addition, we have empirical evidence that the number of anti-patterns in software systems increases over time and only in few cases they are removed through refactoring operations \cite{3}\cite{17}.

All these findings suggest that code smells and anti-pattern need to be carefully detected and
monitored and, whenever necessary, refactoring operations should be planned and performed to deal with them. Unfortunately, the identification and the correction of design flaws in large and non-trivial software systems can be very challenging. These observations call for \textit{planners} capable of supporting the software engineer by (i) identifying anti-patterns, and (ii) designing and applying a refactoring solution to remove them. Such planners, are the focus of the current paper.

\subsection{What is a ``Plan''?}

Our planners   use tables of data with independent features and a dependent class
feature. Classes have weights  that  indicate  what rows are ``bad'' or ``better''. Plans  change   a row such that it is more likely to be ``better''. 
Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
  adjust   feature $Z_j$:

@@

\subsection{From Detection to Planning}\label{sect:p2p}

This paper is about the next step {\em after} detection of bad smells. Suppose developers are presented with  a list of anti-patterns and they do not like what they see. This user may then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''.

Before exploring automatic methods to answer the planning question, we first comment
on two manual methods.

One way to propose changes to a project would be to   ask some smart experienced
person for their opinion on the impact of these anti-patterns and the possible refactoring options. Most advices offered by these developers are based on experiences, sometimes
such advice is an effective strategy and sometimes it is not.

According to Passos et al.~\cite{passos11},  developers
may  assume that the lessons they learn from a few past
projects are general to 
all their future projects. They comment ``past experiences were taken into account without 
much consideration for their context''~\cite{passos11}.  
J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that the 
supposed software engineering  ``Gurus'' rarely use lessons
from past projects to improve their future reasoning and that such poor
past advice can be detrimental to new projects.~\cite{jorgensen09}.
Accordingly, we   propose a ``trust, but verify'' approach.
After a software guru offers some sage wisdom,  it is wise to ask some other oracle 
if there are any better options
(just as a sanity check).
The rest of this paper discusses some methods to build automatic oracles 
to implement that   sanity check.

Another way  to find  changes to a project
might be to rely
on the peer review processes used by the 
SE research community. This approach would propose changes to software
projects that concur with internationally accept best practices. 
There are two
problems with that approach. Firstly, given the rapid pace of change in software
engineering, we may be asking questions for which there is no current widely
accepted concept of
``best practice''.

Secondly, given the diversity of SE products and practices
and personnel, it may well be that the current project being discussed is 
substantively different to prior work. 
Numerous recent {\em local learning} results compare (1) single models
learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
A repeated result in those studies is that the local models generated better effort
and defect predictions (better median results,
lower variance in the predictions). \tion{surprise} of this  paper offers yet another locality result:
\bi
\item
One standard rule in the literature
is that it is useful to implement modules such that they are internally cohesive (use
much of their own local methods) while being loosely coupled with other classes~\cite{Dhama199565}.
\item
While that may be true in general, for particular classes other changes may be more important
(later in this paper, we show one set of results where that is indeed the case).
\ei
In summary, 
it is useful to have automatic methods to recommend changes. Such
methods can fill in for human expertise (if such experts are absent) or 
to offer a second opinion.
Also, prior to making automatic recommendations, it is wise to first stratify the data
(clump it into related examples) then generate advice specific to each clump.
Accordingly, the rest of this paper defines and evaluates
automatic methods to find plans from
  $N$ examples divided  into  many clumps.

%XXX end

%  \begin{figure}[!t]
%  \small
%  \begin{center}
%  \begin{tabular}{r|rr}
%  data set & cases & \% defective\\\hline
%   ant &947& 22\\
%   camel& 1819& 19\\
%  jedit& 1257& 2\\
%  ivy &352& 11\\
%  log4j& 244 &92\\
%  lucebe &442 &59\\
%  poi& 936 &64\\
%  synapse &379 &34\\
%  velocity& 410& 34\\
%  xalan& 2411& 99\\
%  xerces &1055& 74
%  \end{tabular}
%  \end{center}
%  \caption{ Jureczko data: columns in the format of \fig{ck}.}\label{fig:jd}
%  \end{figure}

%  \begin{figure}[!t]
% \scriptsize
% \begin{tabular}{llllll}
%   \hline
%   \rowcolor{lightgray}
% Project & Domain & Lang. & LOC & Features & Config\\\hline
% BDBC: Berkeley DB   & Database & C & 219,811 & 18 & 2560\\
% BDBJ: Berkeley DB   & Database & Java & 42,596 & 32  & 400\\
% Apache & Web Server & C & 230,277 & 9 & 192\\
% SQLite & Database & C & 312,625 & 39 & 3,932,160\\
% LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
% x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
% \end{tabular}

% \caption{Siegmund data.}\label{fig:cpm}
% \end{figure}

% \begin{figure}[!t]
% \includegraphics[width=0.9\linewidth]{figs/BDBC.eps}
% \caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
% \end{figure}

@@

The Jureczko data records the number of known defects for each class using a post-release bug tracking system. The classes are described in terms of nearly two dozen metrics, included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc. For details on the Jureczko data, see  \fig{ck} and \fig{jd}. All the planning methods of this paper reflect on the numeric value of the raw defect counts. The predictor considers defects to be a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise. The choice of this data set was is based on the taxonomy of the smells, in which we focus on the measurable properties needed to detect a given smell. This set of metrics is not restricted and can be easily extended with other metrics.

% The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To obtain the  data, Siegmund et al. perturbed
% the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
% Berkeley database (one written in ``C'' and one in Java). Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref} for the rest)~\cite{sven12}. It's worth noting that in each of the above data sets, several features are interdependent, this is expressed using a \textit{feature model}.
% \fig{bdbc} shows an example of such a feature model defining valid combinations of settings in the Berkley database (``C'' version). These feature models were used by Siegmund et al. to ensure all their perturbations are valid
% (we will use the same models to cull invalid plans).

This section describes XTREE (which we call Method4) and  three  
alternate methods for learning plans.
  XTREE, a novel planner introduced in this work,  uses the decision tree learner of \fig{where}.D.
  The other methods use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}. 
All  methods have the  properties proposed in \tion{prelim}: (1)~they are {\em local learners}
that find   plans   particular to each test case; and (2)~they  are {\em trust-increasing}; 
  i.e. the changed examples are moved {\em closer} to the training data.

Our  description of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma,\omega$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

 {\em Summary1:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be  (which we will call $C_j$).

{\em Assumption1: } Large data sets can be adequately represented by a few dozen (or so) 
centroids.

 {\em Details1:} Method1 clusters project data by reflecting on the independent variables, then
  reports the delta between the cluster centroids. 
  After clustering training data using the WHERE algorithm of \fig{where}.C, Method1
replaces all clusters with a  centroid $C_i$ computed from the median value of each
continuous/discrete feature. Then, it
finds the closest centroid $C_j$ that has a better
performance score. For Jureczko data, ``better'' means fewer defective examples while for the Seigmund data,
``better'' means lower median runtimes for the examples in that cluster.
Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
in $C_j$. 
Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
the nearest centroid $C_i$.  It then proposes a plan for improving that test case
that is the conjunction of all the deltas between $C_i$ and $C_j$.

%\input{tex/pseudo-code-1.tex}

XTREE builds a decision tree,  then generates
plans from the difference between two branches:
(1)~the branch to where you are and (2)~the branch to where you want to be.

One potential problem with Method 1 is the {\em unsupervised} nature of
the clustering algorithm (WHERE) that  
executes without knowledge of the target class.  {\em Supervised} methods assume that it is useful to also reflect on the target class.

XTREE uses a supervised   decision tree algorithm of \fig{where}. D to divide the data.
Next, XTREE builds plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
That code asks three questions, the last of which returns the plan:
\be
\item
What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee

\section{Evaluation Strategy}
\label{sect:eval}
\input{tex/prediction_stats.tex}

Our evaluation strategy (discussed below) divides this data into a training and a testing set.
From the training set we apply a data miner (to learn a quality predictor) and
various planning methods (to learn different plans). Next, we try applying
each of those plans to the  test set and ask the quality predictor to assess the changed examples.
Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
in the changed examples.

As mentioned in the last section,  this approach depends on having effective predictors for assessing the results. It proved to be  more complicated to commission the Jureczko data sets for this study. For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could
be salvaged using the techniques discussed in this section.

%   For the Siegmund data, this criteria was   relatively easy to achieve.
%   The data in those data sets have a continuous class (runtime of the compiled system)
%   so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= (1 - \frac{abs(a - p)}{a})\times100\%$ ({\em higher} values are {\em better}).
% This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
%   x264. 
%   As a preliminary study, we split that data into equal sized train:test groups
%   and trained a Random Forest
%   Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. After repeating this process 40 times, we achieved nearly perfect accuracy of $s=$\{98.45\%, 99.62\%, 94.86\%, 98.89\%, 98.60\%, 99.87\%\};
%   i.e.  we are confident that the predictors from the Siegmund data can assess
% our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)

@@

@@

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=\linewidth]{figs/PercentChanges.png}
\caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
\end{figure*}

>>>>

\usepackage[shortlabels]{enumitem} 
\usepackage{amsmath}
\usepackage{url}
\usepackage{balance}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{ 0eq:#1}}
%\setlist{nolistsep,leftmargin=5mm}
%\usepackage[pdftex]{graphicx}
\usepackage{program}
\newcommand{\Sample}{{\bf SAMPLE}}
\newcommand{\PEEKING}{{\bf PEEKING2}}
%\usepackage[table]{xcolor}
\definecolor{darkgreen}{rgb}{0,0.3,0}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\usepackage{colortbl}
\usepackage{picture}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{listings}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

\title{Learning Which Bad Smells are 
Real Stinkers?}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Tennessee, USA}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract}

@@

When modern software systems are developed, their design 
changes continuously. Due to market constraints, the developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}). These anti-patterns are generally viewed as poor solutions to recurring design problems, possibly deterring the maintainability of the system. In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{khomh12}\cite{li07}\cite{fab15_1}. Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{fab15_1}\cite{chatzigeorgiou10}\cite{arcoverde11}. This makes early detection of these code smells a necessity to maintain sufficient quality in software systems. This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed not only by identifying potential anti-patterns, but also by (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.

Tools for detecting code smells have been the focus of several researchers lately, \cite{Tufano2015}\cite{boussaa13}. There are two  popular techniques for detecting smells: (1) by the making use of static code analysis and code quality metrics\cite{tsantalis09}; (2) Analysing software change history\cite{moha2010}. While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers with planning actions to improve the quality of source code.  Similar concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.

At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. The panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. We call such tools \textit{planners}. In the context of detecting code smells, the planners are required to address the following issues.

@@

In response to this, we propose a novel {\em planning} method called XTREE which reflects on static code quality metrics to learn refactoring options to a software system such that its performance ``improves'' quantitatively. As a measure of source code quality we take into account the number of classes with defects in them. To this end, we use defect count data set by Jureczko et al. for JAVA systems~\cite{jureczko10}. In addition to defects, the aforementioned data set measures a list of static code attributes associated with each class. These code attributes allow us to reason about the measurable properties of anti-patterns.

@@

\bi
\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reducing the number of defects. Specifically, we use XTREE to suggest a number of valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}. In order to ensure the changes are feasible, XTREE generate succinct plans.

\item \textbf{RQ2:} \textit{Can a planner be used to critique bad refactoring options?} In addition to being able to suggesting changes, it's beneficial for planners to be able to quantitatively comment on the quality of certain refactoring options, thereby preventing unnecessary refactoring.

\ei

@@

\section{Motivation}\label{sect:prelim}

The thesis of this paper is that the XTREE tool can stop developers
wasting time on useless software refactorings. Before presenting experiments
that supports that claim, this section motivates the need for 
such tools.

In many fields,  practitioners hang on to persistent beliefs, even
if the evidence for those beliefs is not strong. For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

As in medicine, so too in software engineering.
There is growing evidence that amongst programmers,
many of their  widely-held beliefs are inaccurate and misleading. 
According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects. They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning. They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.

To support the last paragraph, we offer two
specific examples
were  widely-held views in SE that are   now questionable, given recent results. 
Firstly, Nagappan et al. analyzed~\cite{mei15} 
how the {\tt goto} statement was used  in 11,000 Github repositories, 
they found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his famous 1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.

Secondly, in other work with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. Boehm and Basili comment that this phase delay effect
``..has been a major driver in focusing industrial software practice on thorough requirements analysis and design, on early verification and validation, and on up-front prototyping and simulation to avoid costly downstream fixes''~\cite{boehm01}.
After reading the literature, and surveying academics
and practitioners, we can assert that this is a very widely-held belief. Nevertheless,
this effect seems now heavily mitigated by $21^{st}$ software development languages,
tools and development practices.  When we mined the 171 software projects shepherded
by the SEI in the period 2006-2014, we found that phase delay results
in only minor changes in the 
time to resolve issues (usually less than a 30\% increase)~\cite{me16phase}.

These two studies are symptomatic
of a widespread malaise in software engineering.
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  `` (a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

One area where poorly-justifed beliefs can waste valuable project resources is
refactoring in order to remove code bad smells.

\subsection{Need for Planning}
During lifecycle, a software system undergoes continuous changes in order to keep up with the business level demands \cite{lehman79}. Unfortunately, due to market/customers constraints, the source code quality is often neglected with the risk of introducing \textit{code bad smell} (or code-smells) \cite{fowler09}. Code smells are symptomatic of design issues in codes like code understandability and code changeability. If left unaddressed, these can lead to a variety of maintenance problems and faults in software systems \cite{fowler09}\cite{marinescu06}. Fowler \cite{fowler09} and Brown et al. \cite{brown98} defined a catalogue of more than 30 anti-patterns. It is a common notion that the critiques motivated by anti-patterns and code smells would be more comprehensible to developers because many code smells are associated with conventional refactoring operations.

The choice of refactoring operations to address the code smells arise from strongly held beliefs of developers on these topics, and their beliefs in turn are motivated by personal experiences.

The distinction between developers' beliefs and empirical findings is particularly profound in the case of code smells. For instance, \cite{abbes11} showed that the presence of an anti-pattern in the source code does not decrease the developers' performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{abbes11}\cite{yama13}. While the results of this study indicates that single anti-patterns are not harmful, they also reveals that anti-patterns are quite diffuse in software systems and very often code components are affected by more than one anti-pattern. In addition, we have empirical evidence that the number of anti-patterns in software systems increases over time and only in few cases they are removed through refactoring operations \cite{arcoverde11}\cite{chatzigeorgiou10}. Additionally, Yamashita and Moonen\cite{} observed that several developers did not take conscious actions to alleviate bad smells that were present in the code.

Closer inspection revealed that this is not due to the lack of relevance of code smells, but a number of other factors, like: (1) lack of awareness; (2) More importantly, lack of developer friendly tools \cite{}. The study goes on to show that these developers are in need of better support tools that assist them during the software evolution cycle.  Specifically, developers needed tools to (1) be developer friendly; (2) be operable in real-time; and (3) be capable of identifying problematic areas using error prediction.

Therefore to assist developers in their refactoring efforts by providing them with reliable analytics it is useful to have automatic methods that can augment human expertise (or replace them if such experts are absent) in addition to offering a second opinion. Accordingly, the rest of this paper defines and evaluates automatic methods to address the above issues.

\subsection{What is a ``Plan''?}
To address the issues discussed in the previous section, we conjecture that the planners must generate valid and meaningful refactoring operations, we refer to these as \textit{plans}. These plans are synthesized by the planner and are thus unique in that they are agnostic to any prior notions of the developers in addition to them being local to the software project being refactored.

To do this, our planners use tables of data with independent features and a dependent class feature. Classes have weights  that  indicate  what rows are ``bad'' or ``better''. Plans  change   a row such that it is more likely to be ``better''. Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to adjust   feature $Z_j$:

@@

@@

The Jureczko data records the number of known defects for each class using a post-release bug tracking system. The classes are described in terms of nearly two dozen metrics, included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc. For details on the Jureczko data, see  \fig{ck}. All the planning methods of this paper reflect on the numeric value of the raw defect counts to estimate the quality of the class. The predictor considers defects to be a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise. The choice of this data set was is based on the taxonomy of the smells, in which we focus on the measurable properties needed to detect a given smell. This set of metrics is not restricted and can be easily extended with other metrics.

\section{Oracles}
\label{sect:eval}
\input{tex/prediction_stats.tex}

As highlighted previously, there are several detection techniques for anti-patterns, yet these fail to address the issue of checking the proposed changes for their efficacy. Much of the refactoring needs to evaluated for their usefulness. To this end, there needs to be an oracle, a procedure that can be employed to distinguish between good and bad options. The dearth oracles is not unique to this domain, for instance a recent survey by Barr et al \cite{barr15}

Our strategy (discussed in greater detail below) divides this data into a training and a testing set. The purpose of training set is twofold: (1) We apply a data miner to construct an oracle that acts as a quality predictor; (2) We build the planners to reason across the data.

Next, we employ the planners to generate plans to improve the quality of defective classes in the test set. In addition to the plans obtained from our planners we may have some common refactoring options that the developers tend to use. The oracle built previously can be used to assess the the changed examples and this can be used to compare refactoring operations.

It's worth noting that this approach depends on having effective oracles for assessing the results. It proved to be more complicated to commission the Jureczko data sets for this study. For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could be salvaged using the techniques discussed in this section.

@@

\input{tex/algos}

In this we describe the planners used in our study. First, we reintroduce a planner called CD which was an outcome of our previous work~\cite{me12c} which generates plans using a top-down bi-clustering method described in \fig{where}.C to recursively partition the data along the dimension that captures the greatest variability in the data. Then, we present XTREE, a novel planner introduced in this work, which uses a decision tree learner \fig{where}.D for planning.

% This section describes the new XTREE planner and a planner called CD which was an outcome of our previous work~\cite{me12c}. XTREE, a novel planner introduced in this work,  uses the decision tree learner of \fig{where}.D.

Both methods aim to address the issues highlighted in \tion{prelim}. To achieve this: (1)~they find plans, which can be use to generate reliable refactoring options, particular to each test case by reflecting on local data; and (2)~they offer a mechanism for critiquing bad refactoring options.

% (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.

Our  description of the methods adopts the following convention. All variables set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma,\omega$. In this paper, we show our current settings to these variables produces useful results. Elsewhere \cite{krall14,fu:ase15}, we are exploring tuning methods to find better settings but  we have nothing definitive yet to report on auto-tuning planners.

% Method1 computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be (which we will call $C_j$).

Large data sets can be adequately represented by a few dozen (or so) clusters, and meaningful inferences can be drawn from the centroids of these clusters. This method generates plans for improving faulty classes by looking at the difference between the cluster centroid that most closely resembles the defective class from the training data (denoted by $C_i$) and the cluster centroid that represents a better, non-defective, class (denoted by $C_j$).

To do this, CD clusters training data based on the independent variables using the WHERE algorithm of \fig{where}.C and replaces all clusters with their centroids computed from the median value of each continuous/discrete feature. Then, for every cluster $C_i$ it finds the closest centroid $C_j$ that has a better performance score. For Jureczko data, ``better'' means fewer defective examples. CD then caches the delta between the independent features of $C_i$ and $C_j$. For continuous features, this delta is $C_j - C_i$. Finally, for every test case, CD uses the distance measure $d$ shown in \fig{where}.B to find the nearest centroid $C_i$.  It then proposes a plan for improving that test case: the conjunction of all the deltas between $C_i$ and $C_j$.

A potential problem with CD is the {\em unsupervised} nature of the clustering algorithm (WHERE) that executes without knowledge of the target class. It would be beneficial for us to use {\em Supervised} methods that also reflect on the target class.

XTREE was developed to address this problem. It uses a supervised decision tree algorithm of \fig{where}.D to categorize the data Next, XTREE devises plans from the branches of the decision trees using the code of \fig{xtrees_bare}. That code asks three questions, the last of which returns the plan:

\be
\item What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee

In finding the answers to the above questions, XTREE can (a) Generate an wealth of plans (that can be potential refactoring operations); (b) Critique plans based on their location in the decision tree constructed. In addition to the above, XTREE also offers developers a visual medium to validate their proposals for refactoring, see ~\fig{xtree_samp}.

\begin{figure}[btp!]
\centering
\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
\caption{Sample plans from XTREE}\label{fig:xtree_samp}
\end{figure}

@@

\begin{figure}[tb!]
\centering
\includegraphics[width=\linewidth]{figs/thresh.png}
\caption{Thresholds used by DECOR to detect Spaghetti Code~\cite{moha10}.}\label{fig:thresh}
\end{figure}

@@

\begin{figure}[tbp!]
\centering
\includegraphics[width=\linewidth]{figs/PercentChanges.png}
\caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
\end{figure}
